Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-07-10 final/co-regular-EPQ.tex
modify

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-INTRO-GAP-EPQ1L.pdf}
%\vspace{-10pt}
  \caption{An execution that is not $\mathsf{MatchedMaxPriority}^{>}$-linearizable. We represent each operation as a time interval whose left, resp., right, bound corresponds to the call, resp., return action. {\color {blue}Operations adding and removing the same value are aligned vertically.}}
  \label{fig:introduce gap for EPQ1Lar}
\end{figure}

@@

The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains a cycle: $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$. The presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}), {\color {red}as indicated by the following lemma:

@@

%TODO WHY IT IS A SET OF AUTOMATA ? SAY WHAT ARE THE CASES THAT NEED TO BE COVERED. SAY WHAT IS THE CASE WE GIVE IN THE FIGURE. USE $\top$ AS MENTIONED AT THE BEGINNING OF THE SECTION INSTEAD OF 	$c$. USE BIG CAPS FOR SETS OF TRANSITION LABELS.
To define a $\mathsf{MatchedMaxPriority}^>$-complete automaton we need to consider several cases depending on the order between the call/return actions of the $\textit{put}$/$\textit{rm}$ operations that add and respectively, remove the value $b$. For example, the case where the put happens-before the remove (as in \figurename~\ref{fig:introduce gap for EPQ1Lar}) is pictured in \figurename~\ref{fig:automata APQ1Lar-1 in paper}. {\color {red}This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t. the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper} (a)} (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$-s). The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$. {\color {red}In \figurename~\ref{fig:executions APQ1Lar-1 in paper}, we show all the four orders of the call/return actions of adding and removing $b$, and also possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$.} {\color {red}In the long version \cite{CONCUR2017Ahmed}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.} 
%In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.}
{\color {blue}Automata for the other cases can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1 \textwidth]{figures/PIC_AUTO_PQ1Lar-pprr.pdf}
%\vspace{-10pt}
  \caption{Register automaton $\mathcal{A}_{\textit{l-lar}}^1$. We use the following notations: $C_1 = C \cup \{ \textit{ret}(\textit{rm},a) \}$, $C_2 = C \cup \{ \textit{call}(\textit{put},a,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.}
  \label{fig:automata APQ1Lar-1 in paper}
\end{figure}

\begin{figure}[htbp]
  \centering
  %\includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-pprr-in-paper.pdf}
  \includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-fouCase.pdf}
%\vspace{-10pt}
  %\caption{Three possible positions of the first $\textit{call}(\textit{put},a,\_)$ in captured execution}
  \caption{{\color {red}Four cases of ordering actions with value $b$, and possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$}}
  \label{fig:executions APQ1Lar-1 in paper}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6 \textwidth]{figures/PIC-HIS-INTRO-PB-ORDER-EPQ.pdf}
%\vspace{-10pt}
  \caption{An execution that is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^=$.}
  \label{fig:introduce pb order}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8 \textwidth]{figures/PIC-HIS-FiveEnumerations.pdf}
%\vspace{-10pt}
  \caption{{\color {red} Five cases that need to be considered.}}
  \label{fig:five enumerations}
\end{figure}

\begin{figure}[htbp]
  \centering
  %\includegraphics[width=0.8 \textwidth]{figures/PIC-Enumeration-WitnessAutomata.pdf}
  \includegraphics[width=0.8 \textwidth]{figures/PIC-WitnessAutomata-For1.pdf}
%\vspace{-10pt}
  \caption{A case in deriving a $\mathsf{MatchedMaxPriority}^{=}$-complete automaton. We use the following notations: $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true})$, $\textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_1 = C \cup \{ \textit{call}(\textit{put},d,<r) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},d) \}$, $C_4 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},d) \}$.}
  \label{fig:an enumeration and its witness automaton}
\end{figure}

>>>>

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4 \textwidth]{PIC-HIS-INTRO-GAP-EPQ1L.pdf}
%\vspace{-10pt}
  \caption{An execution that is not $\mathsf{MatchedMaxPriority}^{>}$-linearizable. We represent each operation as a time interval whose left, resp., right, bound corresponds to the call, resp., return action. {\color {blue}Operations adding and removing the same value are aligned vertically.}}
  \label{fig:introduce gap for EPQ1Lar}
\end{figure}

@@

The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains a cycle: $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$. The presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable, % (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}), 
{\color {red}as indicated by the following lemma:

@@

%TODO WHY IT IS A SET OF AUTOMATA ? SAY WHAT ARE THE CASES THAT NEED TO BE COVERED. SAY WHAT IS THE CASE WE GIVE IN THE FIGURE. USE $\top$ AS MENTIONED AT THE BEGINNING OF THE SECTION INSTEAD OF 	$c$. USE BIG CAPS FOR SETS OF TRANSITION LABELS.
To define a $\mathsf{MatchedMaxPriority}^>$-complete automaton we need to consider several cases depending on the order between the call/return actions of the $\textit{put}$/$\textit{rm}$ operations that add and respectively, remove the value $b$. For example, {\color {red}the automata for the case} where the put happens-before the remove (as in \figurename~\ref{fig:introduce gap for EPQ1Lar}) is pictured in \figurename~\ref{fig:automata APQ1Lar-1 in paper}. {\color {red}This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t. the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper} (a)} (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$-s). The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_{10} \ldots \rightarrow q_7$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$. {\color {red}In \figurename~\ref{fig:executions APQ1Lar-1 in paper}, we show all the four orders of the call/return actions of adding and removing $b$, and also possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$.} {\color {red}In the long version \cite{CONCUR2017Ahmed}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.}
%In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.}
{\color {blue}Automata for the other cases can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1 \textwidth]{PIC_AUTO_PQ1Lar-pprr.pdf}
%\vspace{-10pt}
  \caption{Register automaton $\mathcal{A}_{\textit{l-lar}}^1$. We use the following notations: $A_1 = A \cup \{ \textit{ret}(\textit{rm},a) \}$, $A_2 = A \cup \{ \textit{call}(\textit{put},a,=r) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $A = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.}
  \label{fig:automata APQ1Lar-1 in paper}
\end{figure}

\begin{figure}[htbp]
  \centering
  %\includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-pprr-in-paper.pdf}
  \includegraphics[width=1 \textwidth]{PIC_HIS_PQ1Lar-fouCase.pdf}
%\vspace{-10pt}
  %\caption{Three possible positions of the first $\textit{call}(\textit{put},a,\_)$ in captured execution}
  \caption{{\color {red}Four cases of ordering actions with value $b$, and possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$}}
  \label{fig:executions APQ1Lar-1 in paper}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-PB-ORDER-EPQ.pdf}
%\vspace{-10pt}
  \caption{An execution that is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^=$.}
  \label{fig:introduce pb order}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8 \textwidth]{PIC-HIS-FiveEnumerations.pdf}
%\vspace{-10pt}
  \caption{{\color {red} Five cases that need to be considered.}}
  \label{fig:five enumerations}
\end{figure}

\begin{figure}[htbp]
  \centering
  %\includegraphics[width=0.8 \textwidth]{figures/PIC-Enumeration-WitnessAutomata.pdf}
  \includegraphics[width=0.8 \textwidth]{PIC-WitnessAutomata-For1.pdf}
%\vspace{-10pt}
  \caption{A case in deriving a $\mathsf{MatchedMaxPriority}^{=}$-complete automaton. We use the following notations: $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true})$, $\textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_1 = C \cup \{ \textit{call}(\textit{put},d,<r) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},d) \}$, $C_4 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},d) \}$.}
  \label{fig:an enumeration and its witness automaton}
\end{figure}
